HMM Model

### 2020-10-22 manuscript/hmm.Rmd
Fixingg issues with HMM models

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care))
  #mutate(best_supportive_care = as.integer(best_supportive_care))

@@

trt_group_m1_rate_data <- rbind(
  tibble(.from = c("Oxygen", "Ventilated"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[2 : length(breathing_s_levels)], .to = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], .to = breathing_s_levels[2 : length(breathing_s_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
         .from = factor(.from, levels = disease_s_levels),
         .to = factor(.to, levels = disease_s_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"))

trt_group_m1_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed)

@@

trt_group_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  trt_group_m1_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_group_m1_fit <- brmhmm(trt_group_m1, cache_file = paste0(cache_dir,"/trt_group_m1.rds"), control = default_control, init = 0.1)

@@

@@

```{r trt_group_m2}
trt_group_m2_formula <- update.formula(trt_group_m1_formula,  ~ . +  (age_norm + sex || rate_group))
trt_group_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m2_fit <- brmhmm(trt_group_m2, cache_file = paste0(cache_dir,"/trt_group_m2.rds"), init = 0.1)

summary(trt_group_m2_fit$brmsfit)
```

@@

```{r trt_group_m3}
trt_group_m3_formula <- update.formula(trt_group_m2_formula,  ~  . + (0 + .rate_id | hospital_id))
trt_group_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m3_fit <- brmhmm(trt_group_m3, cache_file = paste0(cache_dir,"/trt_group_m3.rds"), init = 0.1)

@@

```{r trt_group_m4}
trt_group_m4_formula <-  update.formula(trt_group_m3_formula ~ . + (comorbidities_sum || rate_group)) trt_group_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, trt_group_m1_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m4_fit <- brmhmm(trt_group_m4, cache_file = paste0(cache_dir,"/trt_group_m4.rds"), init = 0.1)

@@

trt_rate_m1_rate_data <- trt_group_m1_rate_data

trt_rate_m1_initial_states <- trt_group_m1_initial_states

@@

```{r trt_rate_m2}
trt_rate_m2_formula <- update.formula(trt_rate_m1_formula,  ~ . +  (age_norm + sex || .rate_id))
trt_rate_m2 <- brmshmmdata(trt_rate_m2_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m2_fit <- brmhmm(trt_rate_m2, cache_file = paste0(cache_dir,"/trt_rate_m2.rds"), init = 0.1)

@@

trt_rate_m3_fit <- brmhmm(trt_rate_m3, cache_file = paste0(cache_dir,"/trt_rate_m3.rds"), init = 0.1)

@@

```{r trt_rate_m4}
trt_rate_m4_formula <-  update.formula(trt_rate_m3_formula ~ . + (comorbidities_sum || .rate_id)) 
trt_rate_m4 <- brmshmmdata(trt_rate_m4_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m4_fit <- brmhmm(trt_rate_m4, cache_file = paste0(cache_dir,"/trt_rate_m4.rds"), init = 0.1)

@@

d_dimer_m1 <- brmshmmdata(d_dimer_m1_formula, d_dimer_serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m1_fit <- brmhmm(trt_m6, cache_file = paste0(cache_dir,"/d_dimer_m1.rds"), init = 0.1)

@@

```{r}
d_dimer_m2_formula <-  update.formula(d_dimer_m1_formula, ~ . + (age_norm + sex + comorbidities_sum || rate_group))

d_dimer_m2 <- brmshmmdata(d_dimer_m2_formula, d_dimer_serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m2_fit <- brmhmm(trt_m6, cache_file = paste0(cache_dir,"/d_dimer_m2.rds"), init = 0.1)

@@

IL_6_m1 <- brmshmmdata(IL_6_initial_states, IL_6_serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m1_fit <- brmhmm(IL_6_m1, cache_file = paste0(cache_dir,"/IL_6_m1.rds"), init = 0.1)

@@

```{r}
IL_6_m2_formula <-  update.formula(IL_6_m1_formula, ~ . + (age_norm + sex + comorbidities_sum || rate_group))

IL_6_m2 <- brmshmmdata(IL_6_initial_states, IL_6_serie_data, 
                  trt_group_m1_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m2_fit <- brmhmm(IL_6_m2, cache_file = paste0(cache_dir,"/IL_6_m2.rds"), init = 0.1)

@@

>>>>

```{r}
serie_data <- wide %>%
  inner_join(data$patient_data, by = c("hospital_id", "patient_id")) %>% 
  mutate(.time = day, .serie = patient_id, .observed = breathing_s) %>%
  mutate(.time = .time +1) %>% 
  filter(.time >= 1) %>%
  group_by(.serie) %>%
  filter(sum(!is.na(.observed)) > 1) %>%
  ungroup() %>%
  #Force explicit dummy coding for the models
  mutate(across(tidyselect::starts_with("took_"), as.integer),
         best_supportive_care = as.integer(best_supportive_care),
         is_male = as.integer(sex == "M")
         )
  #mutate(best_supportive_care = as.integer(best_supportive_care))

@@

simple_rate_data <- rbind(
  tibble(.from = c("Oxygen", "Ventilated"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA", .to = "Discharged", rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[2 : length(breathing_s_levels)], .to = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], rate_group = "improve_one"),
  tibble(.from = breathing_s_levels[1 : (length(breathing_s_levels) - 1)], .to = breathing_s_levels[2 : length(breathing_s_levels)], rate_group = "worsen_one")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
         .from = factor(.from, levels = disease_s_levels),
         .to = factor(.to, levels = disease_s_levels)
  ) %>%
  recode_group("death") %>%
  recode_group("improve_one") %>%
  recode_group("worsen_one") %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"),
         )

@@

trt_group_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  simple_rate_data, simple_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

trt_group_m1_fit <- brmhmm(trt_group_m1, cache_file = paste0(cache_dir,"/trt_group_m1.rds"), control = default_control, init = 0.1, iter = 1500)

@@

```{r, fig.width=12, fig.height=12}
#bayesplot::mcmc_pairs(xx$fit, regex_pars = c("b_.rate_id"))
# bayesplot::mcmc_pairs(trt_group_m1_fit$brmsfit, regex_pars = "worsen_one", np = nuts_params(trt_group_m2_fit$brmsfit))

```{r, fig.width=12, fig.height=12}
# bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, pars = c("b_.rate_idAA_Discharged", "b_.rate_idOxygen_AA"), regex_pars = "improve.*(age|male)")
# 
#bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, regex_pars = c("b_.rate_id"))
```

@@

```{r trt_group_m2}
trt_group_m2_formula <- update.formula(trt_group_m1_formula,  ~ . +  (0 + age_norm + is_male || rate_group))
trt_group_m2 <- brmshmmdata(trt_group_m2_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m2_fit <- brmhmm(trt_group_m2, cache_file = paste0(cache_dir,"/trt_group_m2.rds"), init = 0.1, control = default_control, iter = 1200)

#parnames(trt_group_m2_fit$brmsfit)
summary(trt_group_m2_fit$brmsfit)
```
```{r, fig.width=12, fig.height=12}
# 
#  
# bayesplot::mcmc_pairs(trt_group_m2_fit$brmsfit, regex_pars = "improve_one", np = nuts_params(trt_group_m2_fit$brmsfit))
# 
# 
# bayesplot::mcmc_parcoord(trt_group_m2_fit$brmsfit, regex_pars = "death", np = nuts_params(trt_group_m2_fit$brmsfit))
# bayesplot::mcmc_parcoord(trt_group_m2_fit$brmsfit, regex_pars = "worsen_one", np = nuts_params(trt_group_m2_fit$brmsfit))
```

@@

```{r trt_group_m3}
trt_group_m3_formula <- update.formula(trt_group_m2_formula,  ~  . + (0 + .rate_id | hospital_id))

trt_group_m3 <- brmshmmdata(trt_group_m3_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m3_fit <- brmhmm(trt_group_m3, cache_file = paste0(cache_dir,"/trt_group_m3.rds"), init = 0.1, control = default_control, iter = 1200)

@@

```{r trt_group_m4}
trt_group_m4_formula <-  update.formula(trt_group_m3_formula,  ~ . + (comorbidities_sum || rate_group))

trt_group_m4 <- brmshmmdata(trt_group_m4_formula, serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, simple_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

trt_group_m4_fit <- brmhmm(trt_group_m4, cache_file = paste0(cache_dir,"/trt_group_m4.rds"), init = 0.1, control = default_control, iter = 1200)

@@

trt_rate_m1_rate_data <- simple_rate_data

trt_rate_m1_initial_states <- simple_initial_states

@@

```{r trt_rate_m2}
trt_rate_m2_formula <- update.formula(trt_rate_m1_formula,  ~ . +  (0 + age_norm + is_male || rate_group))
trt_rate_m2 <- brmshmmdata(trt_rate_m2_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m2_fit <- brmhmm(trt_rate_m2, cache_file = paste0(cache_dir,"/trt_rate_m2.rds"), init = 0.1, control = default_control)

@@

trt_rate_m3_fit <- brmhmm(trt_rate_m3, cache_file = paste0(cache_dir,"/trt_rate_m3.rds"), init = 0.1, control = default_control, iter = 1500)

@@

```{r trt_rate_m4}
trt_rate_m4_formula <-  update.formula(trt_rate_m3_formula, ~ . + (comorbidities_sum || .rate_id))

trt_rate_m4 <- brmshmmdata(trt_rate_m4_formula, serie_data, 
                  trt_rate_m1_rate_data, trt_rate_m1_hidden_state_data, trt_rate_m1_initial_states, prior = trt_rate_m1_prior,
                  observed_state_data = trt_rate_m1_observed_state_data)

trt_rate_m4_fit <- brmhmm(trt_rate_m4, cache_file = paste0(cache_dir,"/trt_rate_m4.rds"), init = 0.1, control = default_control, iter = 1500)

@@

d_dimer_m1 <- brmshmmdata(d_dimer_m1_formula, d_dimer_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m1_fit <- brmhmm(d_dimer_m1, cache_file = paste0(cache_dir,"/d_dimer_m1.rds"), init = 0.1, control = default_control, iter = 1500)

@@

```{r}
d_dimer_m2_formula <-  update.formula(d_dimer_m1_formula, ~ . + (0 + age_norm + is_male || rate_group))

d_dimer_m2 <- brmshmmdata(d_dimer_m2_formula, d_dimer_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, d_dimer_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

d_dimer_m2_fit <- brmhmm(d_dimer_m2, cache_file = paste0(cache_dir,"/d_dimer_m2.rds"), init = 0.1, control = default_control, iter = 1500)

@@

IL_6_m1 <- brmshmmdata(IL_6_initial_states, IL_6_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m1_fit <- brmhmm(IL_6_m1, cache_file = paste0(cache_dir,"/IL_6_m1.rds"), init = 0.1, control = default_control, iter = 1500)

@@

```{r}
IL_6_m2_formula <-  update.formula(IL_6_m1_formula, ~ . + (0 + age_norm + is_male || rate_group))

IL_6_m2 <- brmshmmdata(IL_6_initial_states, IL_6_serie_data, 
                  simple_rate_data, trt_group_m1_hidden_state_data, IL_6_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = trt_group_m1_observed_state_data)

IL_6_m2_fit <- brmhmm(IL_6_m2, cache_file = paste0(cache_dir,"/IL_6_m2.rds"), init = 0.1, control = default_control, iter = 1500)

@@

# Complex: Treatments - effects on rate groups

```{r c_trt_m1}
complex_rate_data <- rbind(
  tibble(.from = paste0(c("Oxygen", "Ventilated"), "_worsening"), .to = "Death", rate_group = "death"),
  tibble(.from = "AA_improving", 
         .to = "Discharged", 
         rate_group = "improve_one"),
  tibble(.from = paste0(breathing_s_levels[2 : length(breathing_s_levels)], "_improving"), 
         .to = paste0(breathing_s_levels[1 : (length(breathing_s_levels) - 1)] , "_improving"), 
         rate_group = "improve_one"),
  tibble(.from = paste0(breathing_s_levels[1 : (length(breathing_s_levels) - 1)], "_worsening"),
         .to = paste0(breathing_s_levels[2 : length(breathing_s_levels)], "_worsening"),
         rate_group = "worsen_one"),
  tibble(.from = paste0(breathing_s_levels, "_worsening"),
         .to = paste0(breathing_s_levels, "_improving"),
         rate_group = "to_improving"),
  tibble(.from = paste0(breathing_s_levels, "_improving"),
         .to = paste0(breathing_s_levels, "_worsening"),
         rate_group = "to_worsening")
) %>%
  mutate(.rate_id = factor(paste0(.from, "_", .to), levels = paste0(.from, "_", .to)),
  ) %>%
  mutate(rate_group = fct_relevel(rate_group, "improve_one"))

complex_hidden_state_data <- rbind(
  tibble(base = breathing_s_levels) %>% mutate(corresponding_obs = base) %>% 
    crossing(tibble(state_group = c("_improving", "_worsening"))) %>% 
    mutate(id = factor(paste0(base, state_group))) %>%
    select(-base, -state_group),
  tibble(id = c("Death","Discharged")) %>% mutate(corresponding_obs = id)
)

complex_initial_states <- serie_data %>% filter(.time == 1) %>% arrange(.serie) %>% pull(.observed) %>% paste0(., "_worsening")

c_trt_m1 <- brmshmmdata(trt_group_m1_formula, serie_data, 
                  complex_rate_data, complex_hidden_state_data, complex_initial_states, prior = trt_group_m1_prior,
                  observed_state_data = base_observed_state_data)

c_trt_m1_fit <- brmhmm(c_trt_m1, cache_file = paste0(cache_dir,"/c_trt_m1.rds"), control = default_control, init = 0.1, iter = 1500)

summary(c_trt_m1_fit$brmsfit)

```
```{r}
do_common_pp_checks(c_trt_m1_fit)
```
